---
title: "Data Visualisation for Research"
subtitle: "Part 2: "
author: "Thomas Wise"
output: 
  ioslides_presentation:
  incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Session Objectives and Transferable Skills

<div style="float: left; width: 50%;">
- To embed good practice, through creating replicatable graphics
- To be able to effectively combine complex plots retaining good practice
- To be able to generate publishable plots, in line with common requirements 
</div>
<div style="float: right; width: 50%;">
- Use of replicatable techniques for graph production 
- Use of effective layering techniques in ggplot
- Understanding techniques to make effective publishable plots in line with common standards.
</div>

## Schedule 

- Introduction (5 minutes)
- Review and Recap (5 minutes)
- Part 1: (35 minutes)
    - Advanced Plots (Multiple Layered geom's)
    - Advanced Customization (Legends, Breaks)
- Break (5 minutes)
- Part 2: (35 minutes)
    - Advanced Customization (Colours)
    - Saving, Publishing and Exporting Plots
- Review and Conclusion (5 minutes)


# Last Session Review

# Part 1

## Layering Geoms 

To layer geoms, simply add an additional geom layer to your ggplot() function. If you are using different *mapping = aes()* remember to relocate them to within the appropriate function.

For example: 

```{r, eval = FALSE}

ggplot(data = diamonds, mapping = aes(x, y)) + 
  geom_point()

```

Should become 

```{r, eval = FALSE}

ggplot(data = diamonds) + 
  geom_point(mapping = aes(x, y)) + 
  geom_point(mapping = aes(x, y))

```

## Adding Legends (Colour, discrete)

If you are using a discrete variable, you can use the function: *scale_colour_identity()*. In which you can specify the exact *breaks()*, *labels*, and how to present it *guide = "legend"*. 

```{r, eval = FALSE}

+ scale_colour_identity(name = ??,
                        breaks = c(??, ??),
                        labels = c(??, ??),
                        guide = "legend")

```



## Colour gradients 

If you are using a continuous variable, you can use the function: *scale_colour_gradient()*. In this function you can specify the *low* and *high* points as well as the breaks. 

```{r, eval = FALSE}

+ scale_colour_gradient(low = ??, 
                        high = ??,
                        breaks = ??)

```

## Faceting

Rather than layering plots, you can also present them side by side with *facet_wrap()* or *facet_grid()*. 

You can use it like so; this can be added simply to the end of your ggplot() function. This can be used to specify the rows or columns. 

```{r, eval = FALSE}

+ facet_grid(.~ ??)

```

## Saving and Publishing Plots 

<div style="float: left; width: 50%;">

```{r, eval = FALSE}

ggsave(filename,
       plot = last_plot(),
       height, 
       width, 
       dpi)

```

Easily set the file type by following the *filename* with the format (.jpeg, .png). 

You can also change *last_plot()*, to a gg-variable.

</div>
<div style="float: right; width: 50%;">

- Available formats:
  - JPEG, .jpeg 
  - PNG, .png
  - PDF, .pdf 
  - TIFF, .tiff
  - SVG, .svg
  - EPS, .eps

</div>

## Saving and Publishing Plots 

Alternatively, you can also use the *export* tab in RStudio Plots Panel to export as an *image* or *pdf*. 

This GUI method works in the same function as *ggsave()* but can be easily used if you'd prefer. 

